Pool Model Paper

### 2023-12-07 pool_model.tex
Add model cooperation, corrections to previous Figures

        \dot{T} = \mathcal{F}_{1,2} (R, G_B) \\

@@

    \mathcal{F}_1(R,G)&=&\kappa G_B\\
    \mathcal{F}_2(R,G)&=&\kappa\alpha_B R G_B.

@@

        The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown line) calculated using Eq. (\ref{eq:model_2sp_toxin}).
        The growing bacteria $B$ produces toxin $T$ (a) constantly $\dot{T} = \mathcal{F}_1 = \kappa G_B$ and (b) only during growth process $\dot{T} = \mathcal{F}_2 = \kappa \alpha_B R G_B$ with 

@@

        The toxin $T$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2$.

@@

Model $\mathcal{F}_1$ results in toxin concentration $T_1(t)=\kappa\int_{t_0}^tG(t')dt'$.

@@

On the other hand, model $\mathcal{F}_2$ results in toxin accumulation in steady state $T_2^s=\lim_{t\to\infty}T_2(t)=\kappa(N_t-N_0)$.
To see this we note $\dot{T}_2=\kappa(\dot{L_B}+\dot{G_B})$, which yields: $T_2(t)=\kappa(L_B(t)+G_B(t)-N_{B0})$ where we used $T_2(0)=G_B(0)=0$, $L_B(0)=N_{B0}$.

@@

        The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (a) $\psi$ = 0.5, (b) $\psi$ = 1 or (c) $\psi$ = 2.

@@

@@

Sometimes neighbors, so-called 'social cheaters', can benefit from a certain behavior without expressing it by itself \cite{rainey_evolution_2003}.

@@

\cmt{\cite{rocha_base_2002}: Symbiosis host/bacteria bacterial genomes suffering reductive evolution due to this.
Maybe it's connected with 'social cheaters'?}
\cmt{Production of public goods!}

@@

    \dot{G}_A &=& \mathcal{B}_A(W_B,R)G_A\\
    \dot{G}_B &=& \mathcal{B}_B(W_A,R)G_B\\
    \dot{R} &=&-\mathcal{B}_A(W_B,R)G_A-\mathcal{B}_B(W_A,R)G_B\\
    \dot{W}_A &=& \mathcal{F}_A(R,G_A)\\
    \dot{W}_B &=& \mathcal{F}_B(R,G_B).

@@

$W_A$ and $W_B$ denote the substance produced by species A and B, respectively.
The functions $\mathcal{B}_A(W_B,R)$ and $\mathcal{B}_B(W_A,R)$ capture the resource dependent growth.
Setting $\mathcal{B}_A(W_B,R)=\alpha_A W_BR$, $\mathcal{B}_B(W_A,R)=\alpha_B W_AR$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $W_B$ with $N_t$, $W_A$ with $\kappa_A/\alpha_A$, time with $\alpha_AN_t^2$, defining $\phi=\kappa_A\alpha_B/(\alpha_A^2N_t)$, and $\psi=\kappa_B/(\alpha_A N_t)$ results in:
    \dot{G}_A &=& W_BG_A\left(1 - G_A-G_B\right)\\
    \dot{G}_B &=& \phi W_AG_B\left(1 - G_A-G_B\right)\\
    \dot{W}_A &=& G_A\left(1 - G_A-G_B\right)\\
    \dot{W}_B &=&\psi G_B\left(1 - G_A-G_B\right).

>>>>

        \dot{T}_{B}^{1,2} = \mathcal{F}^{1,2} (R, G_B) \\

@@

    \mathcal{F}^1(R,G)&=&\kappa G_B\\
    \mathcal{F}^2(R,G)&=&\kappa\alpha_B R G_B.

@@

        The two species system competing for the common nutritient $R$ with production of the toxin $T_B$ (brown line) calculated using Eq.~(\ref{eq:model_2sp_toxin}).
        The growing bacteria $B$ produces toxin $T_B$ (a) constantly $\dot{T}_B^1 = \mathcal{F}^1 = \kappa G_B$ or (b) only during growth process $\dot{T}_B^2 = \mathcal{F}^2 = \kappa \alpha_B R G_B$ with 

@@

        The toxin $T_B$ causes the cell death of bacteria $A$ with rate constant $\mu = 0.2$.

@@

Model $\mathcal{F}^1$ results in toxin concentration $T_B^1(t)=\kappa\int_{t_0}^tG(t')dt'$.

@@

On the other hand, model $\mathcal{F}^2$ results in toxin accumulation in steady state $T_B^{2s}=\lim_{t\to\infty}T_B^2(t)=\kappa(N_t-N_0)$.
To see this we note $\dot{T}_B^2=\kappa(\dot{L_B}+\dot{G_B})$, which yields: $T_B^2(t)=\kappa(L_B(t)+G_B(t)-N_{B0})$ where we used $T_B^2(0)=G_B(0)=0$, $L_B(0)=N_{B0}$.

@@

        The relation between inhibition strength of species $B$ and $A$ are defined by control parameter value (a) $\psi$ = 1, (b) $\psi$ = 0.5 or (c) $\psi$ = 2.

@@

It happens if bacteria species as a result of their metabolism produce the compounds that are beneficial for the growth of each other, so-called public goods.

@@

Furthermore, Sometimes neighbors can benefit from a certain behavior without expressing it by itself, so-called 'social cheaters'~\cite{rainey_evolution_2003}.

@@

@@

    \dot{G}_A &=& \mathcal{B}_A(P_B,R)G_A\\
    \dot{G}_B &=& \mathcal{B}_B(P_A,R)G_B\\
    \dot{R} &=&-\mathcal{B}_A(P_B,R)G_A-\mathcal{B}_B(P_A,R)G_B\\
    \dot{P}_A &=& \mathcal{F}_A(R,G_A)\\
    \dot{P}_B &=& \mathcal{F}_B(R,G_B).

@@

$P_A$ and $P_B$ denote the substance produced by species A and B, respectively.
The functions $\mathcal{B}_A(P_B,R)$ and $\mathcal{B}_B(P_A,R)$ capture the resource dependent growth.
Setting $\mathcal{B}_A(P_B,R)=\alpha_A P_BR$, $\mathcal{B}_B(P_A,R)=\alpha_B P_A R$, $\mathcal{F}_A(R,G_A)=\kappa_A RG_A$, $\mathcal{F}_B(R,G_B)=\kappa_B RG_B$, rescaling $G_A$, $G_B$, and $P_B$ with $N_t$, $P_A$ with $\kappa_A/\alpha_A$, time with $\alpha_AN_t^2$, defining $\psi=\kappa_A\alpha_B/(\alpha_A^2N_t)$, and $\phi=\kappa_B/(\alpha_A N_t)$ results in:
    \dot{G}_A = P_B G_A\left(1 - G_A-G_B\right)\\
    \dot{G}_B = \psi P_A G_B\left(1 - G_A-G_B\right)\\
    \dot{P}_A = G_A\left(1 - G_A-G_B\right)\\
    \dot{P}_B =\phi G_B\left(1 - G_A-G_B\right)
    \label{eq:1pool_2sp_cooper}
The solution of the system~\ref{eq:1pool_2sp_cooper} is presented in Figure~\ref{fig:1pool_2sp_cooper}.
Here we can see that if $\psi=1$ and $\phi=1$ (see Figure~\ref{fig:1pool_2sp_cooper}a) the production of the public goods of the two species is happening with the same rate as well as the growth of bacteria.
Increasing the production rate of the public good produced by species $B$ (see Figure~\ref{fig:1pool_2sp_cooper}b) causes increasing the growth of species $A$ that profit from the presence of $B$.
On the other hand, choosing $\psi=2$ (see Figure~\ref{fig:1pool_2sp_cooper}c), we increase the profit that species $B$ has from public goods $P_A$.
That indirectly leads to larger amount of $P_B$ produced by species $B$.
    \includegraphics[width=0.5\textwidth]{Figures/pool_model_2sp_1pool_coop.pdf}
        The two-species system where each of the species release public goods $P$ is defined with equations~(\ref{eq:1pool_2sp_cooper}) and calculated for different parameters:
        (a) $\psi=1, \phi=1$, (b) $\psi=1, \phi=2$ and (b) $\psi=2, \phi=1$.
        Here the red lines describes the evolution of the species $G_A$ and blue lines correspond to species $G_B$.
        Light brown curve corresponds to public good released by species $A$ ($P_A$) and brown curve shows public goods produced by species $B$ ($P_B$).
    \label{fig:1pool_2sp_cooper}
